---
output: 
  github_document:
    pandoc_args: --webtex
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warnings = FALSE,
  collapse = TRUE
)
```

# ymisc

This package contains miscellaneous functions I collected to make simple yet common tasks easier. 


You can install the package from Github via

``` r
remotes::install_github("enixam/ymisc")
```


## General hypothesis testing in multiple linear regression 

A general hypothesis in multiple linear regression can be expressed in the following matrix form 

$$
C\boldsymbol{\beta} = \boldsymbol{d}
$$

where $C$ is an $m \times n$ matrix, $\boldsymbol{\beta}$ is an $p$-dimensional (including the intercept) parameter to estimate, and $\boldsymbol{d}$ is an $m$-dimensional vector. $C$ and $\boldsymbol{d}$ are specified by the hypothesis to be tested. For example, if we want to test the following hypothesis with two constraints for a linear model with 3 predictors 

$$
\begin{aligned}
H_0&: \beta_1 + \beta_3 = 0 \quad \text{and} \quad \beta_2 = 3   \\
H_1&: \text{otherwise} 
\end{aligned}
$$
Then this hypothesis can be expressed as 

$$
\begin{bmatrix}
0 & 1 & 0 & 1 \\
0 & 0 & 1 & 0 \\
\end{bmatrix}
\begin{bmatrix}
\beta_0 \\
\beta_1 \\
\beta_2 \\
\beta_3 
\end{bmatrix}
= 
\begin{bmatrix}
0 \\
3
\end{bmatrix}
$$
`clm` returns the F-statistic and p value associated with such test. It also returns the estimated $\sigma^2$ and $\boldsymbol{\beta}$ under $H_0$, i.e. the equality constrained model. 

```{r}
library(ymisc)
df <- df <- read.table("http://www.stat.ucla.edu/~nchristo/statistics_c173_c273/jura.txt", header=TRUE)
df <- df[, c(-1, -2, -3, -4)]
C <- matrix(
  c(0, 1, 0, 1,
    0, 0, 1, 0),
  nrow = 2, byrow = TRUE)
d <- c(0, 3)

clm_res <- clm(Pb ~ Cd + Co + Cr, data = df, coef_mat = C, d = d)

# F statistic and p-value of 
clm_res$F_stat
clm_res$p_value

# estimate of beta under H_0
clm_res$coefficients
```


In cases of testing significance of individual parameters, t test can be used by setting `t_test = TRUE`. Suppose we want to obtain the t statistic for $\beta_1$ 

```{r}
C <- matrix(c(0, 1, 0, 0), nrow = 1)
d <- 0
clm_t <- clm(Pb ~ Cd + Co + Cr, data = df, coef_mat = C, d = d, t_test = TRUE)

clm_t$t_stat
clm_t$p_value
```

We can verify that the t statistic is the square root of the F statistic, and the p-value is the same. 


```{r}
clm_f <- clm(Pb ~ Cd + Co + Cr, data = df, coef_mat = C, d = d)
sqrt(clm_f$F_stat) # same as t statistic
clm_f$p_value
```

An alternative is using `anova` for nested models, which yields the same result

```{r}
anova(lm(Pb ~  Co + Cr, data = df),
      lm(Pb ~ Cd + Co + Cr, data = df))
```

## Write a data frame groupwise 

There are times when we may wish to write different groups of a data frame to separate files. I used to turn to `puurrr::group_walk` for this, but it is not that easy to control names of such files. `group_write` is a wrapper of`group_walk` and utilizes grouping functions from dplyr to automate this task as well as providing informative files names.   

Suppose we would like to divide passengers in the Titanic according to their age and gender, this would result in 4 groups 

```{r}
path <- tempdir()
titanic <- as.data.frame(Titanic)
titanic <- titanic[rep(1:nrow(titanic), titanic$Freq), -5]
dplyr::count(titanic, Sex, Age)

```

Now use `group_write` to output this 4 groups into different csv files: 

```{r}
group_write(titanic, Sex, Age, dir = path)
list.files(path, "\\.csv$")
```



## Set multiple file encodings at once 

`set_file_enc` uses `iconv` to convert files between encodings. The input can either be a directory or a single file. It works well with import functions from the readr package. 




