---
output: 
  github_document:
    pandoc_args: --webtex=https://ibm.codecogs.com/png.latex?
always_allow_html: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warnings = FALSE,
  collapse = TRUE,
  fig.align = "center"
)
```

# ymisc

<!-- badges: start -->
[![R-CMD-check](https://github.com/enixam/ymisc/workflows/R-CMD-check/badge.svg)](https://github.com/enixam/ymisc/actions)
<!-- badges: end -->

This package contains my personal miscellaneous functions. 

You can install the package from Github via

``` r
remotes::install_github("enixam/ymisc")
```

## Scaffold a `bs4_book` format bookdown project with automatic deployment

```r
use_bookdown("my-awesome-book", tempdir(), gh_action = TRUE)
```

Note that the github action for deploying the book to github pages is designed to use `renv` for installing dependencies. So you'll need to run `renv::init()` in the new project directory before it can work.


## General hypothesis testing in multiple linear regression

<!-- A general hypothesis in multiple linear regression can be expressed in the following matrix form  -->

<!-- $$ -->
<!-- C\boldsymbol{\beta} = \boldsymbol{d} -->
<!-- $$ -->

<!-- where $C$ is an $m \times n$ matrix, $\boldsymbol{\beta}$ is an $p$-dimensional (including the intercept) parameter to estimate, and $\boldsymbol{d}$ is an $m$-dimensional vector. $C$ and $\boldsymbol{d}$ are specified by the hypothesis to be tested. For example, if we want to test the following hypothesis with two constraints for a linear model with 3 predictors  -->

<!-- $$ -->
<!-- \begin{aligned} -->
<!-- H_0&: \beta_1 + \beta_3 = 0 \quad \text{and} \quad \beta_2 = 3   \\ -->
<!-- H_1&: \text{otherwise}  -->
<!-- \end{aligned} -->
<!-- $$ -->
<!-- Then this hypothesis can be expressed as  -->

<!-- $$ -->
<!-- \begin{bmatrix} -->
<!-- 0 & 1 & 0 & 1 \\ -->
<!-- 0 & 0 & 1 & 0 \\ -->
<!-- \end{bmatrix} -->
<!-- \begin{bmatrix} -->
<!-- \beta_0 \\ -->
<!-- \beta_1 \\ -->
<!-- \beta_2 \\ -->
<!-- \beta_3 \\ -->
<!-- \end{bmatrix} -->
<!-- =  -->
<!-- \begin{bmatrix} -->
<!-- 0 \\ -->
<!-- 3 \\ -->
<!-- \end{bmatrix} -->
<!-- $$ -->
<!-- `clm` returns the F-statistic and p value associated with such test. It also returns the estimated $\sigma^2$ and $\boldsymbol{\beta}$ under $H_0$, i.e. the equality constrained model.  -->

```{r}
library(ymisc)
df <- df <- read.table("http://www.stat.ucla.edu/~nchristo/statistics_c173_c273/jura.txt", header=TRUE)
df <- df[, c(-1, -2, -3, -4)]
C <- matrix(
  c(0, 1, 0, 1,
    0, 0, 1, 0),
  nrow = 2, byrow = TRUE)
d <- c(0, 3)

clm_res <- clm(Pb ~ Cd + Co + Cr, data = df, coef_mat = C, d = d)

# F statistic and p-value of
clm_res$F_stat
clm_res$p_value

# estimate of beta under H_0
clm_res$coefficients
```


In cases of testing significance of individual parameters, t test can be used by setting `t_test = TRUE`. Suppose we want to obtain the t statistic for $\beta_1$

```{r}
C <- matrix(c(0, 1, 0, 0), nrow = 1)
d <- 0
clm_t <- clm(Pb ~ Cd + Co + Cr, data = df, coef_mat = C, d = d, t_test = TRUE)

clm_t$t_stat
clm_t$p_value
```

We can verify that the t statistic is the square root of the F statistic, and the p-value is the same.


```{r}
clm_f <- clm(Pb ~ Cd + Co + Cr, data = df, coef_mat = C, d = d)
sqrt(clm_f$F_stat) # same as t statistic
clm_f$p_value
```

An alternative is using `anova` for nested models, which yields the same result

```{r}
anova(lm(Pb ~  Co + Cr, data = df),
      lm(Pb ~ Cd + Co + Cr, data = df))
```

## Write a data frame groupwise

There are times when we may wish to write different groups of a data frame to separate files. I used to turn to `puurrr::group_walk` for this, but it is not that easy to control names of such files. `group_write` is a wrapper of`group_walk` and utilizes grouping functions from dplyr to automate this task as well as providing informative files names.

Suppose we would like to divide passengers in the Titanic according to their age and gender, this would result in 4 groups

```{r}
path <- tempdir()
titanic <- as.data.frame(Titanic)
titanic <- titanic[rep(1:nrow(titanic), titanic$Freq), -5]
dplyr::count(titanic, Sex, Age)
```

Now use `group_write` to output this 4 groups into different csv files:

```{r}
group_write(titanic, Sex, Age, dir = path)
list.files(path, "\\.csv$")
```



## Set multiple file encodings at once

`set_file_enc` uses `iconv` to convert files between encodings. The input can either be a directory or a single file. It works well with import functions from the readr package.



## A light ggplot2 theme

I plan to include more custom ggplot2 themes in this package. Currently there is only one light theme, `theme_clean()`

```{r}
library(ggplot2)
mtcars %>%
  ggplot(aes(wt, mpg, fill = factor(am))) +
  geom_point() +
  geom_smooth() +
  theme_clean()
```

# Convert scripts into markdown code blocks

intended to use with chunk option `results='asis'`

```r
# convert multiple scripts
code_blocks(c(dir1, dir2, dir3))
# convert single script
code_block(file)
```
